---
output: 
    html_document:
      fig_caption: true
      css: custom_styles.css
      includes: 
        in_header: "header_manual.html" 
        after_body: "footer.html"
params:
  network: NETN
  Park: MORR
  state: NJ
  year: 2019
  bcr: 28
  BCRName: Appalachian Mountains
  BBSRegion: S28
  minYear: 2006
  maxYear: 2019
---
      
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# load in packages
library(magrittr)
library(devtools)
library(tidyverse)
library(NCRNbirds)
library(leaflet)
library(cowplot)
library(readxl)
library(lubridate)
library(plotly)
library(knitr)
library(kableExtra)
library(formattable)
library(RColorBrewer)
```

`getParkNames(BirdData[ParkList == params$Park], name.class = "long")`{.tabset .tabset-fade .tabset-pills}
------------------------------------
## {.tabset}
### Overview
<div style="float:right;position:relative;top:10px;padding:5px 5px 5px 10px;margin:0px 5px 10px 5px">
```{r Effmap, echo = FALSE,  fig.height=5.75, fig.width= 5.75, fig.align = 'left', warning= FALSE, comment=FALSE, message=FALSE,}
# Map of AT with HUC10 Shell and AT center line zoomed into selected subregion (NE) showing number of FIA plots per ecoregion as chloropleth. 
ggplot()
```
<p class='capwrap'>
Map of AT with HUC10 Shell and AT center line zoomed into selected subregion (NE) showing number of FIA plots per ecoregion as chloropleth.  </p>
</div>

<h3> Text Block Title </h3>

<p> Text Block
Could use text from Stanke’s website: Forests of the Appalachian National Scenic Trail | rFIA
(Once the basic template is built I can add in the text summarizing the data for the managers.) </p>

<h3>Methods (or other header)</h3>

<p> More text. </p>

### Species Diversity {.tabset }

<p>  Text block describing metrics for most recent Species Diversity. Shannon’s Diversity Index (alpha) Shannon’s Evenness Index (alpha) Species Richness (alpha) e.g., Species diversity | rFIA I would also use this area to describe basic patterns </p> 


#### Shannon’s Diversity Index (alpha)

<p> <b>text in tab in bold</b> </p> 

<p class='caption'>
```{r richness, echo = TRUE, echo=FALSE, fig.height=4, fig.width= 10, fig.align = 'left', fig.cap= "Figure 1. Observed (blue) and estimated (green) species richness over time. Data summaries exclude species not suspected to be breeding in the forest (e.g., flyovers). The number of species estimated (green) is based on the study by Doser et al. (2021), which accounts for imperfect detection and the effect of forest structure. Note that this plot is not showing the total number of bird species in the park because our methods specifically focus on forest songbirds. Other bird species may be present in the park that are not detected during monitoring."}

```
</p>

```{r, richPark, echo=F, fig.align = 'left',fig.cap= "Figure 2. Total number of species detected per park during surveys since monitoring began."}

```

```{r, richTrend, echo=F, fig.cap='Figure 3. Trends in estimated park-level species richness from 2006-2019. Points are the average species richness across all sites, gray regions denote the 95% credible intervals (measure of uncertainty). Inset text is the median (95% credible interval) linear trend estimating change per year in species richness (Doser et al. 2021). Note that observations at ELRO, HOFR, and VAMA were combined under ROVA for the analysis', fig.align='left'}
```


#### Shannon’s Evenness Index (alpha)

<div style="float:right;position:relative;top:10px;padding:5px 5px 5px 10px;margin:0px 5px 10px 5px">
```{r richmap, echo = FALSE,  fig.height=6, fig.width= 6, fig.align = 'left', warning= FALSE, comment=FALSE, message=FALSE}
```
<p class='capwrap'>
Map of cumulative bird species richness detected at `params$Park`'s permanent monitoring sites between `params$minYear` and `params$maxYear`. Click on the sites to see species richness observed per site.</p>
</div>

#### Species Richness (alpha)

### Tree Density and Biomass {.tabset }

<p> text.</p>

<div style="float:right;position:relative;top:10px;padding:1px 1px 1px 1px;margin:0px 5px 10px 5px">
```{r BCI, echo=FALSE, fig.height=4, fig.width= 8, fig.align = 'left',fig.cap= "Figure 4. Annual Bird Community Index values averaged across the park forest. Data summaries exclude species not suspected to be breeding in the forest (e.g., flyovers). The annual BCI value shown is the average (+/-95% C.I.) BCI rating across all monitoring sites. The horizontal axis indicates the year, with the number of points monitored by visit in parenthesis.", message=FALSE}
ggplot()
# Bar chart of mean species density and biomass per species by DBH class size for top 10 most abundant species

# Time series plot showing top 5 species and Fraxinus density/biomass over time per ecoregion. It would be great to show all top 10 species but this could get a bit busy with 10 species X ecoregions). It may not be needed to show ecroregions.

# It would be great to repeat the above plots for total biomass and carbon
```

</div>

### Tree Vital Rates {.tabset}

Text area describing metrics and patterns of tree growth and mortality ( tree basal area, biomass  growth, stand level basal area and biomass growth. Annual mortality and recruitment)

```{r}
# Bar chart showing mean annual growth per top 10 species across region of interest (NE in this example)

# Similar map and bar charts for basal area and biomass growth

# Tabular summary of mean annual mortality rates of top 10 species per ecoregion in area of interest (NE in this example) ----- this one under text not on the right like the others
```


### Regeneration

### Snags

### Downed Woody Debris

### Invasive Plants

### References
